Implementing Distributed Systems Directly with I/O Automata in ioa++

### 2011-09-24 papers/paper1/component_model.tex
New introduction.

Modularity, well-defined interfaces, and composition are essential properties of reusable software.
A unit of software that exhibits all of these properties is called a \emph{component}~\cite{szyperski2002component}.
Modularity implies that a component can be deployed independently of other components.
Well-defined interfaces and interactions allow components to expose their functionality in a regular way that facilitates reasoning about different interactions among them.
Composition allows a group of interacting components to be understood as a cohesive unit.

>>>>

%% Modularity, well-defined interfaces, and composition are essential properties of reusable software.
%% A unit of software that exhibits all of these properties is called a \emph{component}~\cite{szyperski2002component}.
%% Modularity implies that a component can be deployed independently of other components.
%% Well-defined interfaces and interactions allow components to expose their functionality in a regular way that facilitates reasoning about different interactions among them.
%% Composition allows a group of interacting components to be understood as a cohesive unit.

### 2011-09-24 papers/paper1/design.tex
New introduction.

Ioa++ is an implementation in C++ of the component model described in Section~\ref{component_model} for POSIX environments and is available at \url{http://jrwilson.github.com/ioa/}.
Our implementation provides services for executing local actions and facilities for manipulating the set of automata and bindings.
The architecture of our framework is depicted in Figure~\ref{framework_architecture}.
The Scheduler is responsible for selecting and executing local actions subject to the I/O automata model and the concurrency constraints described in Section~\ref{component_model}.
All user automata inherit from a common base class that provides facilities for asynchronously creating and destroying child automata and bindings.
The services provided by the base class are indicated by the Base Automaton layer shown in Figure~\ref{framework_architecture}.
The Base Automaton class contains statically composed System Actions that are used to communicate with the System Automaton.
In most situations, the services provided by the Base Automaton class are too low-level to be used directly by a User Automaton.
An Automaton Manager presents a high-level interface for asynchronously managing a single child automaton using the services of the Base Automaton class.
Similarly, a Binding Manager presents a high-level interface for asynchronously managing a single binding.

>>>>

ioa++ is an implementation in C++ of the component model described in Section~\ref{component_model} for POSIX environments and is available at \url{http://jrwilson.github.com/ioa/}.
Our implementation provides services for executing local actions and facilities for manipulating the set of automata and bindings.
The architecture of our framework is depicted in Figure~\ref{framework_architecture}.
The Scheduler is responsible for selecting and executing local actions subject to the I/O automata model and the concurrency constraints described in Section~\ref{component_model}.
All user automata inherit from a common base class that provides facilities for asynchronously creating and destroying child automata and bindings.
The services provided by the base class are indicated by the Base Automaton layer shown in Figure~\ref{framework_architecture}.
The Base Automaton class contains statically composed System Actions that are used to communicate with the System Automaton.
In most situations, the services provided by the Base Automaton class are too low-level to be used directly by a User Automaton.
An Automaton Manager presents a high-level interface for asynchronously managing a single child automaton using the services of the Base Automaton class.
Similarly, a Binding Manager presents a high-level interface for asynchronously managing a single binding.

### 2011-09-24 papers/paper1/paper.tex
Related work.
New introduction.

\title{Implementing Distributed Systems Directly with I/O Automata in Ioa++}
\author{
  \IEEEauthorblockN{Justin R. Wilson and Christopher D. Gill}
  \IEEEauthorblockA{Department of Computer Science and Engineering\\
    Campus Box 1045, Washington University\\
    One Brookings Drive\\
    St. Louis, MO 63130\\
    {wilsonj,cdgill}@cse.wustl.edu}
  \and
  \IEEEauthorblockN{Guia-Catalin Roman}
  \IEEEauthorblockA{TODO:  Dr. Roman's address}
}

>>>>

\title{Implementing Distributed Systems Directly with I/O Automata in ioa++}
\author{
  \IEEEauthorblockN{Justin R. Wilson and Christopher D. Gill}
  \IEEEauthorblockA{Department of Computer Science and Engineering\\
    Washington University in St. Louis\\
    Campus Box 1045\\
    One Brookings Drive\\
    St. Louis, MO 63130\\
    {wilsonj,cdgill}@cse.wustl.edu}
  \and
  \IEEEauthorblockN{Guia-Catalin Roman}
  \IEEEauthorblockA{School of Engineering\\
    University of New Mexico\\
    Centennial Engineering Center, Suite 3071\\
    1 University of New Mexico, , MSC01 1140\\
    Albuquerque, NM 87131-0001\\
    gcroman@unm.edu}
}

### 2011-09-24 papers/paper1/related_work.tex
New introduction.

Ioa++ can be compared directly to the the IOA language~\cite{garland2003ioa} developed by Lynch et al.
IOA is programming language for I/O automata that can be used with formal verification tools.
Recently, a compiler for IOA has been developed~\cite{tsai2002code},~\cite{tauber2004verifiable},~\cite{tauber2004compiling} and used it to implement a number of protocols~\cite{georgiou2009automated}.
\cite{georgiou2009automated} lists six challenges when developing the compiler: ``Program structuring,'' ``IOA Programs and external services,'' ``Modeling procedure calls,'' ``Composing automata,'' ``Nondeterminism'', and ``Implementing datatypes.''
Programming written in IOA are compiled to Java and executed on hosts communicating using the MPI library.
IOA requires programs to presented in a ``node-channel'' form that allows them to be composed with built in ``mediators'' that implement external services such as MPI.
IOA lacks support for procedure calls which complicates modeling and interfacing with external services; especially when a procedure call blocks.
The IOA compiler composes automata statically using a ``composer'' resulting in a ``node automaton.''
The next action to be executed is determined by a schedule function that resolves the non-determinism in the schedule.

IOA and ioa++ represent two fundamentally different approaches to programming directly with I/O automata.
The development of a new language was necessary for IOA due to their focus on formal verification.
Ioa++ on the other hand uses the existing C++ language and compilers.
IOA encapsulates external services using built in ``mediators'' while ioa++ exposes operating system services via file descriptors relying on the programmer to ensure non-blocking operation.
IOA focuses on static composition to produce a single node program while ioa++ focuses on dynamic composition and allows concurrent execution.
Both IOA and ioa++ require the programmer to add code to schedule actions.

>>>>

ioa++ can be compared directly to the the IOA language~\cite{garland2003ioa} developed by Lynch et al.
IOA is programming language for I/O automata that can be used with formal verification tools.
Recently, a compiler for IOA has been developed~\cite{tsai2002code},~\cite{tauber2004verifiable},~\cite{tauber2004compiling} and used it to implement a number of protocols~\cite{georgiou2009automated}.
\cite{georgiou2009automated} lists six challenges when developing the compiler: ``Program structuring,'' ``IOA Programs and external services,'' ``Modeling procedure calls,'' ``Composing automata,'' ``Nondeterminism'', and ``Implementing datatypes.''
Programming written in IOA are compiled to Java and executed on hosts communicating using the MPI library.
IOA requires programs to presented in a ``node-channel'' form that allows them to be composed with built in ``mediators'' that implement external services such as MPI.
IOA lacks support for procedure calls which complicates modeling and interfacing with external services; especially when a procedure call blocks.
The IOA compiler composes automata statically using a ``composer'' resulting in a ``node automaton.''
The next action to be executed is determined by a schedule function that resolves the non-determinism in the schedule.

IOA and ioa++ represent two fundamentally different approaches to programming directly with I/O automata.
The development of a new language was necessary for IOA due to their focus on formal verification.
ioa++ on the other hand uses the existing C++ language and compilers.
IOA encapsulates external services using built in ``mediators'' while ioa++ exposes operating system services via file descriptors relying on the programmer to ensure non-blocking operation.
IOA focuses on static composition to produce a single node program while ioa++ focuses on dynamic composition and allows concurrent execution.
Both IOA and ioa++ require the programmer to add code to schedule actions.
